### 2019-03-22 paper/sex_selection_spacing-ver2.tex
Edit Data section for grammar

Three factors contribute to the problem here.
First, girls have significantly higher mortality risk than boys.
Second, son preference may increase the probability that parents recall boys 
more readily than girls.
Finally, in NFHS-1 and NFHS-2 enumerators probed only for a missed birth if the
initially reported birth interval was four calendar years or more,
but, given short durations between births, especially after the birth of a girl,
that procedure is unlikely to pick up all missed children.

@@

I focus here on the second through fourth spells, that is, from the second birth until
at most the fourth birth.
The first spell from marriage to the first spell is excluded because of data issues.
In NFHS-1 and NFHS-2 the only information on marriage timing is age at marriage and all
spell durations are therefore imputed.
In NFHS-3 about a third of respondent did not provide complete information on the
timing of marriage and therefore had the length of their first spell imputed using only
the age at marriage.
The result of the imputation is that very large numbers of women end up with very short,
i.e. less than 9 months, or negative spells and that comparing spell duration over time
for the first spell is subject to an excessive amount of noise.

The second and subsequent spells begin nine months after the previous birth 
because that is the earliest we should expect to observe a new birth.
All spells continue until either a child is born or censoring occurs.
Censoring can happen for three reasons:
the survey takes place;
sterilization of the woman or her husband;
or too few births are observed for the method to work.
For all spells censoring is set at 96 months (eight years) after a birth can occur.
With these cut-offs, less than one percent of observed births occur after
the spell cut-off.%
\footnote{
\input{../tables/num_missed.tex}
}

I group spells into four periods based on the year of the previous birth:
1972--1984, 1985--1994, 1995--2004, and 2005--2016.
The periods follow the main changes in availability and legality of 
prenatal sex determination discussed in the Introduction.
The allocation into periods is determined by when conception and therefore 
decisions on sex selection can begin, even if we do not observe
any births until nine months later.

@@

The first variable that is likely to affect the hazard rate in a non-proportional manner
is education.
First, there is evidence, even in the absence of sex selection, that women with
different education levels have different hazard profiles, although the size and direction
of the effect vary across areas and time 
\citep{Tulasidhar1993,Whitworth2002,Bhalotra2008,Kim2010,Soest2018}.
Second, as discussed above, the use of sex selection increases with socioeconomic 
status, especially education, and the use of sex selection is unlikely to be the
same throughout a spell, which means that education cannot enter in a proportional
fashion. \citep{retherford03b,jha06,Portner2015b}.
One potential reason for the correlation between education and sex selection is that
increasing education of mothers correlates closely with lower fertility 
\citep{schultz97}.
This lower fertility leads, in turn, to a higher likelihood of using sex selection
\citep{das_gupta97,Guilmoto2009a,Bongaarts2013,Jayachandran2017}.%
\footnote{
In principle, it is possible that part of the higher likelihood is driven by a stronger
preference for son, independently of desired fertility.
There is, however, evidence that higher education women have shown a decline in
son preference \citep{bhat03,pande07}
}
Another reasons is that having more education is closely associated with higher income and 
wealth, which makes it easier to access illegal services such as prenatal sex 
determination.
It is not possible to separate these two effects here.

>>>>

Three factors contribute to the problem here.
First, girls have a significantly higher mortality risk than boys.
Second, son preference may increase the probability that parents recall boys 
more readily than girls.
Finally, in NFHS-1 and NFHS-2 enumerators probed only for a missed birth if the
initially reported birth interval was four calendar years or more,
but, given short durations between births, especially after the birth of a girl,
that procedure is unlikely to pick up all missed children.

@@

I focus here on the second through fourth spells, that is, on the intervals from the first 
birth until at most the fourth birth.%
\footnote{
I exclude the first spell from marriage to the first spell because of data issues.
In NFHS-1 and NFHS-2 the only information on marriage timing is the age at marriage, and 
all spell durations are therefore imputed.
In NFHS-3 about a third of respondent did not provide complete information on the
timing of marriage and therefore had the length of their first spell imputed using only
the age at marriage.
The result of the imputation is that large numbers of women end up with very short,
i.e., less than nine months, or negative spells and that comparison of spell duration over 
time is subject to an excessive amount of noise.
}
The spells all begin nine months after the previous birth 
because that is the earliest we should expect to observe a new birth, which also means 
that they capture the interpregnancy interval for pregnancies carried to term.
A spell continues until either a child is born or censoring occurs.
Censoring can happen for three reasons:
the survey takes place;
sterilization of the woman or her husband;
or too few births are observed for the method to work.
For all spells censoring is set at 96 months (eight years) after a birth can occur.
With these cut-offs, less than one percent of observed births occur after
the spell cut-off.%
\footnote{
\input{../tables/num_missed.tex}
}

I group spells into four periods based on the year of the previous birth:
1972--1984, 1985--1994, 1995--2004, and 2005--2016.
The periods follow the main changes in the availability and legality of 
prenatal sex determination discussed in the Introduction.
The allocation into periods is determined by when conception and therefore 
decisions on sex selection can begin, even if we do not observe
any births until nine months later.

@@

The first variable that is likely to affect the hazard rate in a non-proportional manner
is education.
First, there is evidence, even in the absence of sex selection, that women with
different education levels have different hazard profiles, although the size and direction
of the effect vary across areas and time 
\citep{Tulasidhar1993,Whitworth2002,Bhalotra2008,Kim2010,Soest2018}.
Second, as discussed above, the use of sex selection increases with socioeconomic 
status, especially education, and the use of sex selection is unlikely to be the
same throughout a spell, which means that education cannot enter proportionally 
\citep{retherford03b,jha06,Portner2015b}.
One potential reason for the correlation between education and sex selection is that
increasing education of mothers correlates closely with lower fertility 
\citep{schultz97}.
This lower fertility leads, in turn, to a higher likelihood of using sex selection
\citep{das_gupta97,Guilmoto2009a,Bongaarts2013,Jayachandran2017}.%
\footnote{
In principle, it is possible that part of the higher likelihood is driven by a stronger
son preference, independently of desired fertility.
There is, however, evidence that higher education women have shown a decline in
son preference \citep{bhat03,pande07}.
}
Another reason is that having more education is closely associated with higher income and 
wealth, which makes it easier to access illegal services such as prenatal sex 
determination.
It is not possible to separate these two effects here.
